---
title: "sc5351_hw3"
author: "Shaoyu Chen"
date: "2023-10-29"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
library(MASS)
df = birthwt
```
##Problem 1

Some medical professionals claim that the average weight of American women is 171 pounds. The column lwt holds the mother’s weight (in pounds) at last menstrual period, i.e. her pre-pregnancy weight. Use this column for the following questions.

Problem 1a

Construct a 95% confidence interval of true mean weight of American women.

n = `r length(pull(birthwt, lwt))`

$\bar{X}$ (sample mean) = `r mean(pull(df, lwt))`

$s$ (sample standard deviation) = `r sd(pull(df, lwt))` 
```{r}
n = length(pull(df, lwt))
x = mean(pull(df, lwt))
s = sd(pull(df, lwt))
```

```{r}
lower_lwt = x - qt(0.975, df = n-1)*(s/sqrt(n))
upper_lwt = x + qt(0.975, df = n-1)*(s/sqrt(n))
```
95% confidence interval is (`r round(lower_lwt, 2)`, `r round(upper_lwt, 2)`).

Problem 1b

Interpret the confidence interval.
 
We are 95% confident that the true population mean of women's weight lies between (`r round(lower_lwt, 2)`, `r round(upper_lwt, 2)`).

Problem 1c

Comment on the validity of the statement above (“Some medical professionals claim that
the average weight of American women is 171 pounds”). In other words, what can we
say about this statement given our confidence interval from part a?

This statement is not valid, because the average weight of American women is 171 pounds is not included in the 95% confidence interval.






##Problem 2

In this data set, we have a variable (smoke) indicating the smoking status of the mothers during pregnancy. Some doctors believe that smoking status is related to weight. Using the columns smoke and lwt, test this claim. (Note: a value of 1 indicates the mother is in the “smoking” group.)

Problem 2a

a) Test for the equality of variances between the two groups. (Use a 5% significance level.)

Test for Equality of Variances

Sample 1: $s_1, ~ n_1, ~ \bar{x}_1$

Sample 2: $s_2, ~ n_2, ~ \bar{x}_2$

Testing the hypotheses: $$  H_0 : \sigma_1^2 = \sigma_2^2 \quad vs \quad H_1 : \sigma_1^2 \ne \sigma_2^2$$

With significance level a pre-specified, compute the test statistic:
$$ F = {\frac{s_1^2}{s_2^2}} \sim F_{n_1-1, n_2-1},under H_0$$

```{r}
var_smoke = 
  birthwt |> 
  filter(smoke == 1) |> 
  pull(lwt) |>
  var()

var_nonsmoke = 
  birthwt |> 
  filter(smoke == 0) |> 
  pull(lwt) |>
  var()

# calculate F statistic
F_test = var_smoke/var_nonsmoke

n_smoke = 
  birthwt |> 
  filter(smoke == 1) |> 
  nrow()
  
n_nonsmoke = 
  birthwt |> 
  filter(smoke == 0) |> 
  nrow()

# calculate critical value  
F_crit = qf(0.975, df1 = 73, df2 = 114)

# decision
ifelse(F_test > F_crit, "reject", "fail to reject")

```
So, it fail to reject

Problem 2b

Given your answer from part a, what kind of hypothesis test will you perform?

I will perform Two-Sample Independent t-test.

Problem 2c

Conduct your chosen hypothesis test from part b at the 10% significance level. What is
your decision regarding the null? Interpret this result in the context of the problem.

Assume the hypotheses: $$H_0 : \mu_1 = \mu_2 \quad vs \quad H_1 : \mu_1 \ne \mu_2$$ 

With significance level α pre-specified, compute the test statistic:
$$t = {\frac{\overline{X_1}-\overline{X_2}}{s\sqrt{\frac{1}{n_1}+\frac{1}{n_2}}}} \sim t_{n_1+n_2-2}$$

Where s is called the pooled sample standard deviation

$$s^2 = \frac{(n_1-1)s_1^2+(n_2-1)s_2^2}{n_1+n_2-2}$$

```{r}
# calculate mean X1,X2
mean_nonsmoke = 
  birthwt |> 
  filter(smoke == 0) |> 
  pull(lwt) |>
  mean()

mean_smoke = 
  birthwt |> 
  filter(smoke == 1) |> 
  pull(lwt) |>
  mean()

# calculate s1,s2
sd_nonsmoke = 
  birthwt |> 
  filter(smoke == 0) |> 
  pull(lwt) |>
  sd()

sd_smoke = 
  birthwt |> 
  filter(smoke == 1) |> 
  pull(lwt) |>
  sd()

# calculate s
s_pool = sqrt(((n_nonsmoke-1)*sd_nonsmoke^2+(n_smoke-1)*sd_smoke^2)/(n_nonsmoke + n_smoke -2))

# calculate T statistic
T_test = (mean_nonsmoke - mean_smoke)/(s_pool*sqrt((1/n_nonsmoke)+(1/n_smoke)))
T_test

# calculate critical value  
T_crit = qt(0.95, n_smoke + n_nonsmoke - 2)
T_crit

# decision
ifelse(T_test > T_crit, "reject", "fail to reject")
```
Based on this result, it fail to reject the null hypothesis, since T_test = 0.6047303 < T_crit = 1.653043. 





##Problem 3 

According to the CDC, approximately 20% of pregnant American women suffer from hypertension. Do our data support this claim? (Use column ht - a value of 1 means the mother is
suffering from hypertension.)

Problem 3a

Conduct a 99% confidence interval and interpret the results. What can we conclude
about the CDC’s claim from this interval?

It is Interval Estimation: One-Sample Proportion

a 100 (1 − a) % confidence interval for one population proportion is given by: $$(\hat{p}-z_{1-a/2}\sqrt{\frac{\hat{p}(1-\hat{p})}{n}},\hat{p}+z_{1-a/2}\sqrt{\frac{\hat{p}(1-\hat{p})}{n}})$$

```{r}
# calculate number of mother is suffering from hypertension
n_ht = 
  birthwt|>
  nrow()

ht_0 = 
  birthwt|>
  filter(ht == 0) |>
  nrow()

ht_1 =
  birthwt|>
  filter(ht == 1) |>
  nrow()

# calculate the value of proportion
ht_prop = ht_1/(ht_1 + ht_0)

# calculate the N(0,1) with a = 0.01
z_ht = qnorm(0.995, 0, 1)

# calculate the lower and upper of CI
lower_ht = ht_prop - z_ht * sqrt(ht_prop*(1 - ht_prop)/n_ht)
upper_ht = ht_prop + z_ht * sqrt(ht_prop*(1 - ht_prop)/n_ht)
```
99% confidence interval is (`r lower_ht`, `r upper_ht`).

We are 99% confident that the population proportion of pregnant American women suffer from hypertension lies between (`r lower_ht`, `r upper_ht`). Based on this result, the claim that CDC has approximately 20% of pregnant American women suffer from hypertension is not correct, because 0.2 does not include in 99% confidence interval.



Problem 3b

Conduct a one-sided hypothesis test at the a = 0.1 level. In this test, we want to see if
the true proportion is indeed less than the claimed 20%. What can we conclude about the CDC’s claim?

I will use One-Sample Test for Binomial Proportion,tests for One-Population Proportion, Normal Theory Methods

Assume the hypothesis: $$H_0 : p = p_0 \quad vs \quad H_1 : p < p_0$$

With significance level α pre-specified, compute the test statistic:
$$ z = {\frac{\hat{p} - p_0}{\sqrt{p_0(1 - p_0)/n}}}\sim N(0,1)$$

```{r}
# calculate critical value  
Z_crit = qnorm(0.95, 0, 1)
Z_crit

# calculate Z statistic
Z_test = (ht_prop-0.2)/sqrt(0.2*(1-0.2)/n_ht)
Z_test

# decision
ifelse(abs(Z_test) > Z_crit, "reject", "fail to reject")
```
Based on this result, |Z_test| = 4.691685 > Z_crit = 1.644854, so it reject the null hypothesis, which means the true proportion is indeed less than the claimed 20%. for CDC's claim, approximately 20% of pregnant American women suffer from hypertension is correct.





##Problem 4

Is there a difference between uterine irritability in the group of pregnant women who smoke vs the group of pregnant women that don’t smoke? (Use columns ui and smoke.)
Conduct a hypothesis test at the a = 0.01 level. What can we conclude about the proportions
of women with uterine irritability between the smoking groups?

First I will use Tests for Two-Population Proportions, Normal Theory Methods:

Assume the hypothesis: $$H_0 : p_1 = p_2 \quad vs \quad H_1 : p_1 \ne p_2$$

The test statistic is given by:
$$z = {\frac{\hat{p_1} - \hat{p_2}}{\sqrt{\hat{p}\hat{q}(\frac{1}{n_1}+\frac{1}{n_2})}}}\sim N(0,1)$$ and where $$\hat{p} = \frac{n_1\hat{p_1}+n_2\hat{p_2}}{n_1+n_2}$$

```{r}
# calculate p1, p2 and p
smoke_ui = 
  birthwt|>
  filter(smoke == 1) |> 
  group_by(ui) |>
  filter(ui == 1) |>
  nrow()

nonsmoke_ui = 
  birthwt|>
  filter(smoke == 0) |> 
  group_by(ui) |>
  filter(ui == 1) |>
  nrow()

# number of nonsmoke = n1, number of smoke = n2
n1 = 
  birthwt |>
  filter(smoke == 0) |>
  nrow()

n2 = 
  birthwt |>
  filter(smoke == 1) |>
  nrow()

# calculate p1 and p2 
p1 = nonsmoke_ui / n1
p2 = smoke_ui / n2

# calculate p
p = (n1*p1+n2*p2) / (n1 + n2)

# calculate Z statistic
Z_test2 = (p1 - p2) / sqrt(p*(1-p)*((1/n1)+(1/n2)))
Z_test2

# calculate critical value 
Z_crit2 = qnorm(0.995, 0, 1)
Z_crit2

# decision
ifelse(abs(Z_test2) > Z_crit2, "reject", "fail to reject")
```
According to this result, At 0.01 significance level, we can’t reject the null hypothesis, because |Z_test2| = 0.8545449 < Z_crit2 = 2.575829. we can say that there is not a significant difference between proportions of women with uterine irritability between the smoking groups.




##Problem 5

Is race related to birth weight? (Use columns race and bwt.)

```{r}
race_df =
  birthwt |> 
  group_by(race) |> 
  count(race)
race_df
```


Problem 5a: 

What test would be most appropriate to answer this question?

ANOVA test.

Problem 5b

What assumptions are we making by using this test? Are all assumptions met?

1. There are k populations of interest (k > 2), 
2. The samples are drawn independently from the underlying populations;
3. Homoscedasticity: the variances of the k populations are equal;
4. Normality: the distributions of the error terms are normal (could be
relaxed when the sample size is large)
They are all met!!! k = 3, the variances of the k populations are equal, the distributions of the error terms are normal, and the samples are drawn independently from the underlying populations.

Problem 5c

Conduct the test at the 5% significance level and interpret your results. Be sure to write
the hypotheses you are testing.

Assume the hypothesis:

$$H_0 : \mu_1 = \mu_2 = \mu_3 \quad vs \quad H_1 : at least two means are not equal$$

Compute the test statistic:

$$F = {\frac{BetweenSS/(k-1)}{WithinSS/(n-k)}}\sim F_{k-1,n-k}$$

distribution under $H_0$

Rejection rules: Reject $H_0$: if F > $F_{k-1,n-k,1-a}$

Fail to reject $H_0$: if F $\leqslant F_{k-1,n-k,1-a}$




Problem 5d

Perform multiple comparisons - which races are significantly different? Interpret your
results


```{r}
pairwise.t.test(birthwt$bwt,birthwt$race,p.adj="bonferroni")
  
```
Based on this result, race 1 is different.



